Assessing the Global COVID19 Impact on Air Transport with Open Data

### 2021-06-27 paper-old-failed-compile.Rmd
some small add on, had to convert to MS Word as pdf kept failing. Not sure what goes wrong.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(stats)
library(zoo)
```

```{r globalstuff, message=FALSE}
## set bookdown specs/defaults for high quality output
#---------- check for the settings

@@

@@

```{r}
#quick peek
#Read "x0_data-preparation.Rmd" first, then
full_dataset

temp <- full_dataset %>% group_by(DATE, ADEP_REG, ADES_REG) %>% summarize(FLIGHTS = n(), .groups = "drop") %>%
  filter(ADEP_REG %in% c("BR", "EU"), ADES_REG  %in% c("BR", "EU")) %>%
  mutate(ROUTE = paste(ADEP_REG, ADES_REG, sep = "-"), .keep = "unused", .before = "FLIGHTS") %>%
  pivot_wider(names_from = ROUTE, values_from = FLIGHTS) 
temp

#First Look
temp1 <- full_dataset %>% group_by(DATE, ADEP_REG, ADES_REG) %>% summarize(FLIGHTS = n()) %>% mutate(ROUTE = paste(ADEP_REG, ADES_REG, sep = "-"), .keep = "unused")
temp1

temp1 %>% ggplot(aes(x = DATE)) +
  geom_line(aes(y = (FLIGHTS), color = ROUTE))

#First stab at different levels of traffic

n <- 0.5
temp1 %>% ggplot(aes(x = DATE)) +
  geom_point(aes(y = `EU-EU`, color = `EU-EU` > quantile(`EU-EU`[month(DATE) %in% 1:2], probs = n)), shape = 4) +
  geom_point(aes(y = `BR-BR`, color = `BR-BR` > quantile(`BR-BR`[month(DATE) %in% 1:2], probs = n)), shape = 1) +
  labs(y  = "Flights") + 
  theme(legend.position = "bottom")

#Normalized by the median of the last 3 months of the dataset
temp1 %>% pivot_longer(cols = 2:5, names_to = "ROUTE", values_to = "FLIGHTS") %>% group_by(ROUTE) %>%
  mutate(MOVING_MEDIAN = quantile(FLIGHTS[month(DATE) %in% month(last(DATE))-2:month(last(DATE))], probs = 0.5), NORM_FLTS = FLIGHTS/MOVING_MEDIAN)
temp2

@@

```{r}

1.2.2

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(stats)
library(zoo)
```

```{r globalstuff, message=FALSE}
## set bookdown specs/defaults for high quality output
knitr::opts_chunk$set(
   echo = FALSE, message = FALSE, warning = FALSE    # messages
  , dpi = 320               #, dev.args = list(type = "cairo") #
  , dev = "ragg_png"        # works better than cairo
  , fig.retina = 2      # resolution and device
#  , fig.align = "center", fig.show = "hold"          # positioning of figure
#  , fig.width = 6, fig.asp = 0.618
#  , out.width = "85%"                                # outside dim of figure
                      )

@@

## load study data - as long as not outsourced to data prep
fns <- list.files(path = "./data/", pattern = "base_osn_", full.names = TRUE )
col_spec <- cols(
  ADEP = col_character(),
  ADEP_CTRY = col_character(),
  ADEP_REG = col_character(),
  ADES = col_character(),
  ADES_CTRY = col_character(),
  ADES_REG = col_character(),
  TYPE = col_character(),
  DATE = col_date(format = ""),
  CALL = col_character()
)
ds  <- fns %>% map_dfr(.f = ~ read_csv(.x))
```

```{r eval=FALSE}
## CHECK WHAT CAN BE REMOVED FROM HERE AND IS NOW OUTSOURCED TO DATA PREP

@@

```{r}
#quick peek
#Read "x0_data-preparation.Rmd" first, then
ds

temp <- ds %>% group_by(DATE, ADEP_REG, ADES_REG) %>% summarize(FLIGHTS = n(), .groups = "drop") %>%
  filter(ADEP_REG %in% c("BR", "EU"), ADES_REG  %in% c("BR", "EU")) %>%
  mutate(ROUTE = paste(ADEP_REG, ADES_REG, sep = "-"), .keep = "unused", .before = "FLIGHTS") %>%
  pivot_wider(names_from = ROUTE, values_from = FLIGHTS) 
head(temp)

#First Look
temp1 <- ds %>% group_by(DATE, ADEP_REG, ADES_REG) %>% summarize(FLIGHTS = n()) %>% mutate(ROUTE = paste(ADEP_REG, ADES_REG, sep = "-"), .keep = "unused")
head(temp1)

temp1 %>% ggplot(aes(x = DATE)) +
  geom_line(aes(y = (FLIGHTS), color = ROUTE))
```

<!-- NEEDED TO INAVTIVATE THE FOLLOWING - I COULD NOT RECONSTRUCT IT -->

```{r eval=FALSE}
#First stab at different levels of traffic

n <- 0.5
temp1 %>% ggplot(aes(x = DATE)) +
  geom_point(aes(y = `EU-EU`, color = `EU-EU` > quantile(`EU-EU`[month(DATE) %in% 1:2], probs = n)), shape = 4) +
  geom_point(aes(y = `BR-BR`, color = `BR-BR` > quantile(`BR-BR`[month(DATE) %in% 1:2], probs = n)), shape = 1) +
  labs(y  = "Flights") + 
  theme(legend.position = "bottom")
```

<!-- needed to deactivate the following:
* threw error with column 5
* could not reconcile which operation would have created the 5th col
-->

```{r, eval=FALSE}
#Normalized by the median of the last 3 months of the dataset
temp1 %>% pivot_longer(cols = 2:5, names_to = "ROUTE", values_to = "FLIGHTS") %>%   
  group_by(ROUTE) %>%
  mutate(MOVING_MEDIAN = quantile(FLIGHTS[month(DATE) %in% month(last(DATE))-2:month(last(DATE))], probs = 0.5), NORM_FLTS = FLIGHTS/MOVING_MEDIAN)

head(temp2)

@@
